RVD2: An ultra-sensitive variant detection model for low-depth targeted next-generation sequencing data

### 2013-11-26 doc/bioinformatics2013/RVD2.tex
Update manuscript

The HCC1187 cell line is derived from  epithelial cells from primary breast tissue from a 41 y/o adult with TNM stage IIA primary ductal carcinoma. A matched normal cells were derived from lymphoblastoid cells from whole blood. 
% TODO: check matched cells line origin
% TODO: mention tumor purity here

@@

Figure~\ref{fig:MAF} shows the posterior mean and 95\% credible intervals for $\mu_j$ for called variant positions with $\bar{n} = 41,573$ and MAF = XX\%. All of the called positions show a clear difference between the case and control error rates. The posterior mean estimates are all shrunken towards the global error rate parameter $\mu_0 = 0.0023$ due to the hierarchical structure of the model.

\begin{figure}[h]
\begin{center}
\includegraphics[width=120mm]{pdf_figs/Synthetic_MuBarPlot.pdf}
\caption{Estimated minor allele fraction for called variants in 0.1\% dilution.}
\label{fig:MAF}
\end{center}
\end{figure}
% from folder '2013-11-24_synthetic_mu_barplot_optimalT_Qsd_0_1'

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=160mm]{pdf_figs/comparison_table_ss.pdf}
\caption{Sensitivity/Specificity comparison of RVD2 with other variant calling algorithms.}
\label{tbl:comparison_ss}
\end{center}
\end{figure}
% from folder 'rvd2\results\2013-09-19_operating_characteristics'

@@

Figure~\ref{fig:brca_MAF} shows the estimated minor allele frequencies for the normal and tumor samples at the called locations. Positions chr7:743899X>X, chr7:753635X>X and chr7:780960X>X are germline homozygous mutations. Positions chr7:754371 and chr7:758813 are called heterozygous in the normal sample. In the tumor cells we identified all five mutations chr7:743899X>X, chr7:753635X>X, chr7:754371X>X, chr7:758813X>X, and chr7:780960X>X. Positions chr7:754371X>X and chr7:758813X>X are identified as loss-of-heterozygosity events.

% TODO: replace X>X with the reference and mutated base here
\begin{figure}[h]
\begin{center}
\includegraphics[width=120mm]{pdf_figs/HCC1187_MuBarPlot.pdf}
\caption{Estimated minor allele fraction for called variants in PAXIP1 gene.}
\label{fig:brca_MAF}
\end{center}
\end{figure}
%('rvd2\results\2013-11-22_HCC1187_PAXIP1_genome_Qsd_0_1')

The original research describing this sample used Strelka to identify mutations in the same sample. They identified XXX and XXX as variant, but missed the other three variants.

@@

% from folder '2013-11-24_depthM_plot_avg10_Qsd_0_1', dilution rate is 0.1.

>>>>

The HCC1187 cell line is derived from  epithelial cells from primary breast tissue from a 41 y/o adult with TNM stage IIA primary ductal carcinoma. A matched normal cells were derived from lymphoblastoid cells from whole blood. 
% TODO: check matched cells line origin
% double checked by Yuting, didn't find the matched cells line origin
% TODO: mention tumor purity here

@@

Figure~\ref{fig:MAF} shows the posterior mean and 95\% credible intervals for $\mu_j$ for called variant positions with $\bar{n} = 5584$ and MAF = 1.0\%. All of the called positions show a clear difference between the case and control error rates. The posterior mean estimates are all shrunken towards the global error rate parameter $\mu_0 = 0.0023$ due to the hierarchical structure of the model.

\begin{figure}[h]
\begin{center}
\includegraphics[width=120mm]{pdf_figs/Synthetic_Mubarplot_MAF1_0_Depth5584.pdf}
\caption{Estimated minor allele fraction for called variants in 1.0\% dilution.}
\label{fig:MAF}
\end{center}
\end{figure}
% from folder '2013-11-15_synthetic_mu_barplot_optimalT', dilution rate is 0.1, downsampling rate is 0.01.

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=160mm]{pdf_figs/comparison_table_ss.pdf}
\caption{Sensitivity/Specificity comparison of RVD2 with other variant calling algorithms.}
\label{tbl:comparison_ss}
\end{center}
\end{figure}
% from folder 'rvd2\results\2013-09-19_operating_characteristics'
% rvd2 data are from folder '2013-11-12_one_replicate_optT_with_chi','2013-11-12_one_replicate_T_zero_with_chi2','2013-11-15_six_replicates_optT_with_chi2','2013-11-15_six_replicates_T_zero_with_chi2'

@@

Figure~\ref{fig:brca_MAF} shows the estimated minor allele frequencies for the normal and tumor samples at the called locations. Positions chr7:743899XC$\rightarrow$T, chr7:753635T$\rightarrow$C and chr7:780960C$\rightarrow$T are germline homozygous mutations. Positions chr7:754371 and chr7:758813 are called heterozygous in the normal sample. In the tumor cells we identified all five mutations chr7:743899C$\rightarrow$T, chr7:753635T$\rightarrow$C, chr7:754371T$\rightarrow$C, chr7:758813G$\rightarrow$A, and chr7:780960C$\rightarrow$T. Positions chr7:754371T$\rightarrow$C and chr7:758813G$\rightarrow$A are identified as loss-of-heterozygosity events.

% TODO: replace X>X with the reference and mutated base here
% Done

\begin{figure}[h]
\begin{center}
\includegraphics[width=120mm]{pdf_figs/HCC1187_MuBarPlot.pdf}
\caption{Estimated minor allele fraction for called variants in PAXIP1 gene.}
\label{fig:brca_MAF}
\end{center}
\end{figure}
%('rvd2\results\2013-11-21_HCC1187_PAXIP1_mu_barplot_T0')

The original research describing this sample used Strelka to identify mutations in the same sample. They identified chr7:154760439 as variant, but missed the other four variants.

@@

% from folder '22013-11-15_depthM_plot_avg10', dilution rate is 0.1.
